Solving Collective Spin Hamiltonians with Decoherence

### 2018-09-29 soc_squeezing/cumulant_expansions.tex
simplified some expressions in the appendix on changing operator bases

We wish to change our basis for all exchange-symmetric operators on a
system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
\begin{align}
  \S_{\v m} \equiv \sum_{\p{\v j}} P_{\v m\v j}
\end{align}
where
\begin{align}
  \v m \equiv \p{m_+,m_\z,m_-},
  &&
  \v j \equiv \p{j_1^+,j_2^+,\cdots j_{m_+}^+,
    j_1^\z,j_2^\z,\cdots j_{m_\z}^\z,
    j_1^-,j_2^-,\cdots j_{m_-}^-},
\end{align}
and for shorthand we define
\begin{align}
  \sum_{\p{\v j}} X_{\v j} \equiv
  \sum_{j_a^\mu=j_b^\nu \iff \p{a,\mu}=\p{b,\nu}} X_{\v j},
  &&
  P_{\v m\v j} \equiv
  \prod_\mu \prod_{a=1}^{m_\mu}
  \sigma_\mu^{\p{j_a^\mu}}.
\end{align}
In words, $\S_{\v m}$ is a $\p{m_++m_\z+m_-}$-spin operator with
$m_\mu$ single-spin operators of type $\sigma_\mu$ addressing the
spins indexed by $j_1^\mu,j_2^\mu,\cdots j_{m_\mu}^\mu$.  We now
expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$ into
terms that have exactly $s$ spins addressed by both operators:
\begin{align}
  \S_{\v m} \S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k,s}} P_{\v m\v j} P_{\v n\v k},
  &&
  \sum_{\p{\v j,\v k,s}} X_{\v j\v k} \equiv
  \sum_{\substack{\p{\v j},\p{\v k} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}}
  X_{\v j\v k},
  \label{eq:SS_PP}
\end{align}
where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
in $P_{\v m\v j}$ address the same spin as a $\sigma_\nu$ operator in
$P_{\v n\v k}$, we have
\begin{align}
  \S_{\v m}\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v m,\v n,\v r,\v J,\v K}}
  f_{\v m\v n\v r} Q_{\v r\v J} \tilde Q_{\v r\v K},
  \label{eq:SS_PQ}
\end{align}
where $\set{J_a^\mu}$ and $\set{K_a^{\mu\nu}}$ are respectively the
indices of spins addressed by one and two single-spin operators in
each term of \eqref{eq:SS_PP}; the sum over
$\p{\v m,\v n,\v r,\v J,\v K}$ implies a sum over all possible values
of $\v r,\v J,\v K$ with the restrictions
\begin{align}
  \sum_{\mu,\nu} r_{\mu\nu} = s,
  &&
  0 \le \sum_\nu r_{\mu\nu} \le m_\mu,
  &&
  0 \le \sum_\mu r_{\mu\nu} \le n_\nu,
\end{align}
\begin{align}
  \abs{\set{J_a^\mu}\cup\set{K_a^{\mu\nu}}}
  = \sum_\mu\p{m_\mu+n_\mu} - s;
\end{align}
the factor $f_{\v m\v n\v r}$ accounts for the number of ways to pair
operators in $P_{\v m\v j}$ and $p_{\v n\v k}$ as prescribed by
$\v r$, or
\begin{align}
  f_{\v m\v n\v r} \equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu\nu} r_{\mu\nu}!},
\end{align}
and
\begin{align}
  Q_{\v r\v J} \equiv \prod_\mu
  \prod_{a=1}^{\abs{\set{J_b^\mu}}} \sigma_\mu^{(J_a^\mu)},
  &&
  \tilde Q_{\v r\v K} \equiv \prod_{\mu,\nu}
  \prod_{a=1}^{\abs{\set{K_b^{\mu\nu}}}}
  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
\end{align}

>>>>

We wish to change our basis for all exchange-symmetric operators on a
system of $N$ qubits from $S_+^\ell S_\z^m S_-^n$ to
\begin{align}
  \S_{\v m} \equiv \sum_{\p{\v j}} P_{\v m\v j}
\end{align}
where
\begin{align}
  \v m \equiv \p{m_+,m_\z,m_-},
  &&
  \v j \equiv \p{j_1^+,j_2^+,\cdots j_{m_+}^+,
    j_1^\z,j_2^\z,\cdots j_{m_\z}^\z,
    j_1^-,j_2^-,\cdots j_{m_-}^-},
\end{align}
and for shorthand we define
\begin{align}
  \sum_{\p{\v j}} X_{\v j} \equiv
  \sum_{j_a^\mu=j_b^\nu \iff \p{a,\mu}=\p{b,\nu}} X_{\v j},
  &&
  P_{\v m\v j} \equiv \prod_{\mu,a} \sigma_\mu^{\p{j_a^\mu}}.
\end{align}
In words, $\S_{\v m}$ is a $\p{m_++m_\z+m_-}$-spin operator with
$m_\mu$ single-spin operators of type $\sigma_\mu$ addressing the
spins indexed by $j_1^\mu,j_2^\mu,\cdots j_{m_\mu}^\mu$.  We now
expand a product of basis operators $\S_{\v m}$ and $\S_{\v n}$ into
terms that have exactly $s$ spins addressed by both operators:
\begin{align}
  \S_{\v m} \S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k,s}} P_{\v m\v j} P_{\v n\v k},
  &&
  \sum_{\p{\v j,\v k,s}} X_{\v j\v k} \equiv
  \sum_{\substack{\p{\v j},\p{\v k} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}}
  X_{\v j\v k},
  \label{eq:SS_PP}
\end{align}
where the sum over $s$ terminates after $s = \sum_\mu\p{m_\mu+n_\mu}$.
Collecting terms in which $r_{\mu\nu}$ of the $\sigma_\mu$ operators
in $P_{\v m\v j}$ address the same spin as a $\sigma_\nu$ operator in
$P_{\v n\v k}$, we have
\begin{align}
  \S_{\v m}\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v m,\v n,\v r,\v J,\v K}}
  f_{\v m\v n\v r} Q_{\v r\v J} \tilde Q_{\v r\v K},
  \label{eq:SS_PQ}
\end{align}
where $\set{J_a^\mu}$ and $\set{K_a^{\mu\nu}}$ are respectively the
indices of spins addressed by one and two single-spin operators in
each term of \eqref{eq:SS_PP}; the sum over
$\p{\v m,\v n,\v r,\v J,\v K}$ implies a sum over all possible values
of $\v r,\v J,\v K$ with the restrictions
\begin{align}
  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
  &&
  \sum_{\mu,\nu} r_{\mu\nu} = s,
  &&
  \sum_\nu r_{\mu\nu} \le m_\mu,
  &&
  \sum_\mu r_{\mu\nu} \le n_\nu,
\end{align}
\begin{align}
  \abs{\set{J_a^\mu}} = \sum_\mu\p{m_\mu+n_\mu} - 2 s,
  &&
  \abs{\set{K_a^{\mu\nu}}} = s,
  &&
  \set{J_a^\mu} \cap \set{K_a^{\mu\nu}} = \set{};
\end{align}
the factor $f_{\v m\v n\v r}$ accounts for the number of ways to pair
operators in $P_{\v m\v j}$ and $P_{\v n\v k}$ as prescribed by
$\v r$, or
\begin{align}
  f_{\v m\v n\v r}
  &\equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\rho>\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho>\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
  &= \sp{\prod_\mu \p{m_\mu}_{\sum_\rho r_{\mu\rho}}
    \p{n_\mu}_{\sum_\rho r_{\rho\mu}}}
  \sp{\prod_{\mu\nu} r_{\mu\nu}!}^{-1},
\end{align}
where $\p{n}_k\equiv\prod_{j=0}^{k-1}\p{n-j}$ is a falling factorial;
and finally
\begin{align}
  Q_{\v r\v J} \equiv \prod_{\mu,a} \sigma_\mu^{(J_a^\mu)},
  &&
  \tilde Q_{\v r\v K} \equiv \prod_{\mu,\nu,a}
  \sigma_\mu^{(K_a^{\mu\nu})} \sigma_\nu^{(K_a^{\mu\nu})}.
\end{align}
